### 2025-01-15 reviewer-response.md

*The manuscript makes too much use of unfamiliar terminology taken from
earlier papers without defining terms.  For example, see line 233 and
Reference [18].  See also the paragraph at lines 306 to 310. All terms
used must be precisely defined in the manuscript to allow proper
assessment of the manuscript.*

We have addressed this by making the paper more self-contained.

*The Results section is a slog to read and parts of the Discussion seem
redundant.  Take lines 651 to 668 as an example.  Tightening up the
Results and the Discussion can only help the exposition.*

offload to figure captions where possible, don't spell out everything (TODO MORGAN)

### 2025-01-16 reviewer-response.md

*The description of the method is clear, although the description of the run time
costs is ambiguous. Overall, CAKES is an interesting contribution to the field
of NN-search.*

# Major comments

*The big O notation by default suggest worst-case run time complexity. In this
manuscript, it seems that complexity is a mixture of expected run time
complexity and under the loose assumption that the data has particular
characteristics (e.g., low fractal dimension, the tree is balanced). These
assumptions should be made clear in the intro and when giving the run time
expressions.*

We agree that because the time complexity is highly dependent on the manifold structure of the data, this was insufficiently rigorous. We have attempted to address this in a few ways. First, we added the condition "when the dataset has a manifold structure" to our statement about asymptotic complexity of entropy-scaling search in Section 1.2. Additionally, we added a paragraph to Section 1.2 justifying our approach to complexity analyses, and restated it in Section 2.2.2 and Section 2.2.5 before giving runtime expressions.

@@

*For example, line 215, does the cost of partitioning of O(n log(n)) depend on
the tree being balanced? It is an important hypothesis because line 197 states
that "the imbalance in the tree is a feature, not a bug, as it reflects the
underlying structure of the data". That may look like a contradiction.*

We have added Section 4.4 to compare the performance of balanced and unbalanced clustering.
We show that balanced trees perform strictly worse during search.

*The cost on line 260 uses $\hat{r}$, which is a mean radius. Now that cost is in
expectation (in addition to the balanced tree hypothesis).*

This is correct; we added justification for our approach to complexity analyses in Sections 1.2, 2.2.2, and 2.2.5.

@@

*The manuscript makes too much use of unfamiliar terminology taken from
earlier papers without defining terms.  For example, see line 233 and
Reference [18].  See also the paragraph at lines 306 to 310. All terms
used must be precisely defined in the manuscript to allow proper
assessment of the manuscript.*

We have addressed this by making the paper more self-contained.

*The Results section is a slog to read and parts of the Discussion seem
redundant.  Take lines 651 to 668 as an example.  Tightening up the
Results and the Discussion can only help the exposition.*

We attempted to address this by omitting unnecessary details, offloading observations to figure captions, and eliminating redundant sections of the discussion.

### 2025-01-16 sections/1-introduction.tex
more offloading to figure captions

The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality.
In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a hierarchical clustering approach.
CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language.

>>>>

The entropy-scaling search paradigm exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms exhibit asymptotic complexity that scales with topological properties of the dataset, instead of its cardinality, when the dataset has a manifold structure.
In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a hierarchical clustering approach.
CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language.

We also provide a theoretical analysis of the time complexity of CAKES's algorithms in Sections~\ref{sec:methods:knn-search:repeated-rnn-complexity} and~\ref{sec:methods:knn-search:complexity-of-sieve-methods}. 
These analyses are not worst-case analyses in the traditional sense, as they do not assume the worst possible dataset, namely, a uniform distribution.
Given that CAKES's algorithms are intended to be used on datasets with a manifold structure, complexity analysis assuming a uniform distribution of data would be uninformative. 
As a result, we assume in our analyses that the dataset has a manifold structure, as quantified by low fractal dimension and metric entropy.
For our purposes, it makes more sense to analyze the expected performance of CAKES's algorithms under these assumptions.

### 2025-01-16 sections/2-methods.tex
more offloading to figure captions

@@

\begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
        \mathcal{O}
        \Bigg(
            \underbrace{
                \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
            }_{\textrm{tree-search}}
            \ + \
            \underbrace{
                \overbrace{k}^{\textrm{output size}} \cdot
                \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
            }_{\textrm{leaf-search}}
        \Bigg)
        \label{eq:methods:repeated-rnn-complexity1}
    \end{gather}
    where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
\end{theorem}

@@

Due to their similarity, we combine the complexity analyses of both Sieve methods.
For these methods we again use the terminology of tree-search and leaf-search.
Tree-search navigates the cluster tree and adds clusters to $Q$.
Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.

\begin{theorem}
    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
        \label{eq:methods:breadth-first-sieve-complexity}
    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( T \log T + L \log k \Big).
        \label{eq:methods:depth-first-sieve-complexity}
    \end{equation}
\end{theorem}

>>>>

As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:rnn-complexity} assumes that the dataset $X$ has a manifold structure.

@@

\begin{theorem} Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Then time complexity of performing Repeated $\rho$-NN search on $X$ with query $q$ is \begin{gather}
        \mathcal{O}
        \Bigg(
            \underbrace{
                \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
            }_{\textrm{tree-search}}
            \ + \
            \underbrace{
                \overbrace{k}^{\textrm{output size}} \cdot
                \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
            }_{\textrm{leaf-search}}
        \Bigg)
        \label{eq:methods:repeated-rnn-complexity1}
    \end{gather}
    where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of the dataset, $d$ is the LFD of the dataset, and $k$ is the number of nearest neighbors.
    \label{thm:methods:rnn-complexity}
\end{theorem}

@@

Due to their similarity, we combine the complexity analyses of both Sieve methods.
For these methods we again use the terminology of tree-search and leaf-search.
Tree-search navigates the cluster tree and adds clusters to $Q$.
Leaf-search exhaustively searches \textit{some} of the clusters in $Q$ to find the $k$ nearest neighbors.
As stated in Section~\ref{sec:intoduction:entropy-scaling-search}, Theorem~\ref{thm:methods:sieve-complexity} assumes that the dataset $X$ has a manifold structure.

\begin{theorem}
    Let $X$ be a dataset and $q$ a query sampled from the same distribution (i.e., arising from the same generative process) as $X$. Let $T \coloneqq \mathcal{O} \big( \lceil d \rceil \cdot \log \mathcal{N}_{\hat{r}}(X) \big)$ and   $L \coloneqq \mathcal{O} \left( k \cdot \bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d \right)$, where $\mathcal{N}_{\hat{r}}(X)$ is the metric entropy of $X$, $d$ is the LFD of $X$, $\hat{|C|}$ is the mean cardinality of clusters overlapping the query ball, and $k$ is the number of nearest neighbors. Then, for dataset $X$ and query $q$, the time complexity of performing Breadth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( (T + L ) \log (T + L ) \Big)
        \label{eq:methods:breadth-first-sieve-complexity}
    \end{equation} and the the time complexity of performing Depth-First Sieve search is \begin{equation}
        \mathcal{O} \Big( T \log T + L \log k \Big).
        \label{eq:methods:depth-first-sieve-complexity}
    \end{equation}
    \label{thm:methods:sieve-complexity}
\end{theorem}
